---
title: "pyNET: Pattern assessment (ODD)"
editor: visual
format: 
  html:
    toc: true
    number-sections: true
    code-fold: true
knitr:
  opts_chunk:
    out.width: "100%"
execute: 
  cache: false
  warning: false
---

```{r}
library(tidyverse)
library(data.table)
library(kableExtra)

library(ggpubr)

ggplot <- function(...){
  ggplot2::ggplot(...) + 
        scale_color_viridis_d(end = 0.8) +
        scale_fill_viridis_d(end = 0.8)
}
    
theme_set(theme_bw())
```

```{r}
envs = c(0:2)
spes = c(0:2)
grafts = c(0:3)
```

```{r}
out_str = c("tdens", "dbh_mean", "age_mean", "root_mean")
out_net = c("fgraft", "gdens", "gsm", "gsx", "fred")
out_gro = c("gs", "dist_mean", "nlinks")
```

```{r}
stands = fread(file = "../data/stands_50.txt") %>%
    data.frame(.) %>%
    mutate_at(vars(ROOTING, RESOURCES, GRAFTING), as.factor) %>%
    rename(
        "gsm" = "gs_mean",
        "gsx" = "gs_max",
        "fred" = "p_non_red_no2",
        "fred_all" = "p_non_red"
    )

groups = fread(file = "../data/groups_50.txt") %>%
    data.frame(.) %>%
    select(-gIDpy) %>% distinct(.) %>%
    rename("gs" = "groupSize",
           "nlinks" = "noLinks") 
```

# Patterns

-   Pattern 1: Root graft frequency decreases with tree density.
-   Pattern 2: Group density increases with tree density.
-   Pattern 3: Group size decreases with tree density.
-   Pattern 4: The network is scale free (i.e., node degree follows a power law distribution).
-   Pattern 5: The majority of groups have a non-redundant shape.

# P1-P3: Spearman Correlation

```{r}
p1p3_spearman = stands %>%
    gather(., key, value, fgraft, gdens, gsm) %>%
    group_by(GRAFTING, ROOTING, RESOURCES, key) %>%
    reframe(rxy = cor.test(value, tdens)$estimate,
            p.value = cor.test(value, tdens)$p.value) %>%
    mutate(signif = ifelse(p.value < 0.05, "*", "ns"),
           pattern = factor(key,
                            labels = c("P1", "P2", "P3"))) 
```

```{r}
p1p3_spearman %>%
    ggplot(., aes(x = key, y = rxy)) +
    geom_hline(yintercept = 0, col = "grey") +
    geom_boxplot() +
    geom_point(aes(alpha = signif, shape = signif),
               position = position_jitterdodge()) +
    scale_alpha_manual(values = c(1, 0.2)) +
    facet_wrap( ~ pattern, scales = "free_x") +
    labs(x = "Y",
         y = "Spearman's rxy")
```

```{r}
p1p3_spearman %>%
    ggplot(., aes(x = ROOTING, 
                  col = GRAFTING,
                  # alpha = GRAFTING,
                  y = rxy)) +
    geom_hline(yintercept = 0, col = "grey") +
    geom_boxplot() +
    geom_point(aes(alpha = signif, shape = signif),
               position = position_jitterdodge()) +
    scale_alpha_manual(values = c(1, 0.2)) +
    facet_wrap( ~ pattern, scales = "free_x") +
    labs(#x = "Y",
         y = "Spearman's rxy") +
    facet_wrap(~key)
```

```{r}
p1p3_spearman %>% 
    mutate(c = ifelse(signif == "*", 1, 0),
           pos = ifelse(rxy > 0, 1, 0)) %>% #View()
    group_by(key) %>% 
    reframe(p_sign = sum(c)/n()*100,
            p_pos = sum(pos)/n()*100,
            rxy_m = mean(rxy),
            rxy_sd = sd(rxy)) %>% 
    mutate_if(is.numeric, round, 1) %>% #View()
   kbl(caption = "Summary of Spearman correlations across scenarios. 'p_sign' indicates the % of significant scenarios and 'p_pos' indicates the % of positive correlations.") %>%
   kable_styling(full_width = F)
```

# P4: The network is scale free (i.e., node degree follows a power law distribution).

```{r}
nd_distribution = fread(file = "../data/nd_distribution.txt") %>%
    data.frame(.) %>%
    mutate(ROOTING = factor(SPE),
           GRAFTING = factor(GRAFT),
           RESOURCES = factor(ENV)) %>% 
   filter(RESOURCES %in% envs)  %>% 
   filter(GRAFTING %in% grafts)  %>% 
   filter(ROOTING %in% spes) 

nd_power_law = fread(file = "../data/nd_power_law.txt") %>%
    data.frame(.) %>%
    mutate(ROOTING = factor(SPE),
           RESOURCES = factor(ENV),
           GRAFTING = factor(GRAFT)) %>%
    filter(RESOURCES %in% envs) %>%
    filter(GRAFTING %in% grafts) %>%
    filter(ROOTING %in% spes) %>%
    mutate(signif = ifelse(KS.p < 0.05, "*", "ns"),
           lab.y = as.numeric(ROOTING)/4+0.15)
```

```{r}
nd_distribution %>% 
    group_by(ROOTING, GRAFTING, RESOURCES, k) %>%
    mutate(p_k_m = mean(p_k)) %>% #View()
    ggplot(., aes(x = k, y = p_k, col = ROOTING)) +
    geom_line(size = 0.3, alpha = 0.3, aes(group = setup)) +
    geom_point(aes(y = p_k_m), size = 2) +
    geom_text(
        nd_power_law,
        mapping = aes(
            x = 7,
            y = lab.y,
            label = signif,
            col = ROOTING
        ),
        show.legend = FALSE
    ) +
    scale_x_continuous(breaks = seq(1, max(nd_distribution$k) + 1, 3)) +
    facet_grid(RESOURCES ~ GRAFTING, labeller = label_both) +
    labs(x = "Node degree")
```

```{r}
nd_power_law %>% 
    group_by(signif) %>% 
    reframe(p = round(n()/nrow(nd_power_law)*100, 1),
            gsm = mean(gsm),
            gsx = mean(gsx))
```

# P5: The majority of groups have a non-redundant shape.

```{r}
stands %>% 
    filter(fred != 0) %>% 
    gather(., key, value, fred_all, fred) %>% 
    ggplot(., aes(x = ROOTING, fill = GRAFTING,
                  alpha = RESOURCES, y = value)) +
    geom_hline(yintercept = 50, col = "grey") +
    geom_boxplot() +
    guides(alpha = guide_legend(override.aes = aes(fill = "grey"))) +
    facet_wrap(~key, ncol = 1)
```

- Percentage of scenarios with fred > 50 %

```{r}
bind_rows(
    stands %>%
        group_by(ROOTING, GRAFTING, RESOURCES) %>%
        reframe(fred = mean(fred)) %>%
        ungroup() %>%
        mutate(no_scen = n()) %>%
        filter(fred > 50) %>%
        reframe(p_sce_fred = n() / no_scen * 100) %>%
        distinct() %>% mutate(groups = "> 2 trees"),
    stands %>%
        group_by(ROOTING, GRAFTING, RESOURCES) %>%
        reframe(fred = mean(fred_all)) %>%
        ungroup() %>%
        mutate(no_scen = n()) %>%
        filter(fred > 50) %>%
        reframe(p_sce_fred = n() / no_scen * 100) %>%
        distinct()%>% mutate(groups = "all")
)

```
# Simultainous occurrence

```{r}

matches = bind_rows(
    p1p3_spearman %>%
        mutate(
            match = case_when(
                pattern %in% c("P1", "P3") & rxy < 0 ~ "match",
                pattern %in% c("P2") & rxy > 0 ~ "match",
                .default = "fail"
            )
        ) %>%
        select(GRAFTING:RESOURCES, pattern, match),
    nd_power_law %>%
        mutate(
            match = ifelse(signif == "*", "match", "match_ns"),
            pattern = "P4"
        ) %>%
        select(GRAFTING, ROOTING, RESOURCES, pattern, match),
    stands %>%
        group_by(ROOTING, GRAFTING, RESOURCES) %>%
        reframe(fred = mean(fred_all)) %>%
        mutate(
            match = ifelse(fred > 50, "match", "fail"),
            pattern = "P5"
        ) %>%
        select(GRAFTING, ROOTING, RESOURCES, pattern, match)
) 

matches = matches %>%
    spread(., pattern, match) %>%
    unite("allP", P1:P5, remove = F) %>%
    unite("notP3", P1, P2, P4, P5, remove = F) %>%
    mutate(
        match_sig = ifelse(str_detect(allP, "fail") |
            str_detect(allP, "ns"),
        "fail",
        "match"),
        match_all = ifelse(str_detect(allP, "fail"),
        "fail",
        "match"),
        match_all_but_P3 = ifelse(str_detect(notP3, "fail"),
        "fail",
        "match")
        )
```
- All Patterns

```{r}
matches %>%
    gather(., key, value, match_all, match_sig, 
           match_all_but_P3) %>% 
    group_by(key, value) %>% 
    reframe(n = n()) %>% 
    spread(., value, n) %>% 
    mutate(tot = fail + match)
```

- Only Pattern P3

```{r}
matches %>% 
    group_by(P3) %>% 
    reframe(n())
```


```{r}
matches %>% 
    gather(., key, value, P1:P5) %>% 
    mutate(value = factor(value,
                          levels = c("match", "match_ns",
                                     "fail"),
                          labels = c("succeed", 
                                     "succeed but P4 \nnon-significant",
                                     "fail"))) %>% 
    ggplot(., aes(y = GRAFTING,
                  x = key,
                  fill = value)) +
    geom_tile(col = "darkgrey", alpha = 0.5) +
    coord_fixed() +
    scale_fill_viridis_d(end = 0.8, direction = -1) +
    facet_grid(ROOTING~RESOURCES, labeller = label_both) +
    labs(x = "Pattern",
         fill = "Reproduction") +
    theme_classic()
ggsave("pattern_matched_2.jpg", width = 6, height = 4)
```


```{r}
RES.labs <- c("RESOURCES 0:\nfixed", 
              "RESOURCES 1:\nvariable", 
              "RESOURCES 2:\nseasonal & variable")
names(RES.labs) <- c("0", "1", "2")

matches %>%
    mutate(counts = str_count(allP, "match"),
           counts_ns = counts - str_count(allP, "ns"),
           lab = ifelse(counts != counts_ns, 
                        paste(counts, "*"),
                        counts)) %>%
    # filter(counts != counts_ns) %>% 
    # View()
    ggplot(., aes(x = GRAFTING, y = ROOTING, fill = counts)) +
    geom_tile(col = "white", alpha = 0.3) +
    scale_fill_viridis_c(end = 0.8, direction = -1) +
    geom_text(aes(label = lab)) +
    guides(fill = F) +
    coord_fixed() +
    facet_wrap( ~ RESOURCES, 
                labeller = labeller(RESOURCES = RES.labs)) +
    theme_classic()


```
```{r}
ggsave("patterns_matched.jpg", width = 8, height = 3., dpi = 400)
```


# Summary

## 1 Scenario

```{r}
RO = 0 # large
RE = 0 # fixed
GR = 3 # costs&exchange
```

```{r}
sss = stands %>%
    filter(ROOTING == RO & RESOURCES == RE & GRAFTING == GR) %>% distinct()

nd_d = nd_distribution %>%
    filter(ROOTING == RO & RESOURCES == RE & GRAFTING == GR)

nd_pl = nd_power_law %>%
    filter(ROOTING == RO & RESOURCES == RE & GRAFTING == GR)
```

```{r}
labs = list("tdens"=expression(paste(Tree~density~(trees~ha^-1))),
            "gdens"=expression(paste(Group~density~(groups~ha^-1))),
            "fgraft"="Grafting frequency (%)",
            "gsm"="Mean group size",
            "fred"="Non-redundant groups (%)")
```

```{r}
fig_p1 = sss %>% 
    ggplot(., aes(x = tdens, y = fgraft)) +
    geom_point(alpha = 0.7) +
    stat_cor(method = "spearman", digits = 1) +
    labs(x = labs["tdens"][[1]],
         y = labs["fgraft"][[1]])

fig_p2 = sss %>% 
    ggplot(., aes(x = tdens, y = gdens)) +
    geom_point(alpha = 0.7) +
    stat_cor(method = "spearman", digits = 1) +
    labs(x = labs["tdens"][[1]],
         y = labs["gdens"][[1]])

fig_p3 = sss %>% 
    ggplot(., aes(x = tdens, y = gsm)) +
    geom_point(alpha = 0.7) +
    stat_cor(method = "spearman", digits = 1) +
    labs(x = labs["tdens"][[1]],
         y = labs["gsm"][[1]])

fig_p4 = nd_d %>%
    group_by(ROOTING, k) %>%
    mutate(p_k_m = mean(p_k)) %>%
    ggplot(., aes(x = k, y = p_k)) +
    geom_line(size = 0.3, alpha = 0.3,
              aes(group = setup)) +
    geom_point(aes(y = p_k_m), size = 2) +
    scale_x_continuous(breaks = seq(1, max(nd_d$k) + 1, 3)) +
    labs(x = "Node degree",
         y = "P(k)")

fig_p5 = sss %>% 
    filter(fred != 0) %>%
    ggplot(., aes(x = "", y = fred)) +
    geom_hline(yintercept = 50, col = "grey") +
    geom_boxplot() +
    geom_jitter() +
    guides(alpha = guide_legend(override.aes = aes(fill = "grey"))) +
    labs(x = "",
         y = labs["fred"][[1]])


ggarrange(fig_p1, fig_p2, fig_p3, fig_p4, fig_p5,
          labels = c("(a)", "(b)", "(c)", "(d)", "(e)"))

```

```{r}
ggsave("fig_8.jpg", width = 9, height = 5.5, dpi = 400)
```

